<link rel="stylesheet" href="http://netdna.bootstrapcdn.com/bootstrap/3.0.3/css/bootstrap.min.css">
<style type="text/css"> body {padding: 10px 30px 10px 30px;} table,th, td {text-align: center;}</style>
<style>
td.tableRow
{
text-align:center;
}
</style>


Market Segmentation for a B2B Company
========================================================

**INSEAD Data Analytics Group Project**


Some Code/Analysis from the course material/tools
---------------------------------------------------------

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
#let's remove the first 2 rows and make the first row into column_names
#ProjectData = tail(ProjectData, -3)
#let's code the "SURVEY_VALUE_CHAIN" column first
#non_numeric_row = ProjectData[,"V3"]
#thevalues = unique(non_numeric_row)
#num_values = rep(0,nrow(ProjectData))
#for (iter in 1:length(thevalues))
#  num_values[which(non_numeric_row == thevalues[iter])]<- iter
#ProjectData[,"V3"] <- num_values

factor_attributes_used = unique(sapply(factor_attributes_used,function(i) min(ncol(ProjectData), max(i,1))))
ProjectDataFactor=ProjectData[,factor_attributes_used]

# let's make the data into data.matrix classes now so that we can easier visualize them
ProjectDataFactor = data.matrix(ProjectDataFactor)
ProjectData = data.matrix(ProjectData)
```

This is how the first `r min(max_data_report, nrow(ProjectData))` data looks (ignoring the NON-numerical columns):
<br>

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectDataFactor,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>


<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectDataFactor),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>


```{r, results='asis'}
ProjectDatafactor_scaled=apply(ProjectDataFactor,2, function(r) {if (sd(r)!=0) res=(r-mean(r))/sd(r) else res=0*r; res})
```



```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
thecor = round(cor(ProjectDataFactor),2)
colnames(thecor)<-colnames(ProjectDataFactor)
rownames(thecor)<-colnames(ProjectDataFactor)
## printing the result in a clean-slate table
cat(renderHeatmapX(thecor, border=1, center = 0,vrange_up = 1, vrange_down = 1))
```


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Component",1:ncol(UnRotated_Factors),sep=" ")
```

```{r echo=FALSE, comment=NA, warning=FALSE, error=FALSE,message=FALSE,results='asis'}
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table


row=1:nrow(Variance_Explained_Table)
name<-paste("Component No:",row,sep="")
Variance_Explained_Table<-cbind(name,Variance_Explained_Table)
Variance_Explained_Table<-as.data.frame(Variance_Explained_Table)
colnames(Variance_Explained_Table)<-c("Components", "Eigenvalue", "Percentage_of_explained_variance", "Cumulative_percentage_of_explained_variance")

m<-gvisTable(Variance_Explained_Table,options=list(width=1200, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'),formats=list(Eigenvalue="#.##",Percentage_of_explained_variance="#.##",Cumulative_percentage_of_explained_variance="#.##"))
print(m,'chart')
```
<br> <br>


```{r Fig1, echo=FALSE, comment=NA, results='asis', message=FALSE, fig.align='center', fig=TRUE}

eigenvalues  <- Variance_Explained_Table[,2]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
Line         <- gvisLineChart(as.data.frame(df), xvar="components", yvar=c("eigenvalues","abline"), options=list(title='Scree plot', legend="right", width=900, height=600, hAxis="{title:'Number of Components', titleTextStyle:{color:'black'}}", vAxes="[{title:'Eigenvalues'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(Line, 'chart')
```



```{r echo=FALSE, comment=NA, warning=FALSE,message=FALSE,results='asis'}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used

```


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}

Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Component",1:ncol(Rotated_Factors),sep=" ")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

show_data <- Rotated_Factors 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
<br> <br>


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)

show_data <- Rotated_Factors_thres 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
<br> <br>


<div class="row">
<div class="col-md-6">

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
NEW_ProjectData <- round(Rotated_Results$scores[,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste("Derived Variable (Factor)",1:ncol(NEW_ProjectData),sep=" ")

if (factors_selected >=2){ 
  
  show_data <- as.data.frame(NEW_ProjectData) 
  show_data = show_data[1:min(max_data_report,nrow(show_data)),]
  row<-rownames(show_data)
  dfnew<-cbind(row,show_data)
  change<-colnames(dfnew)
  change[1]<-"Observation"
  colnames (dfnew)<-change
  m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
  print(m1,'chart')
  
  } else {
    print(xtable(NEW_ProjectData, caption="Only 1 derived variable generated", digits=3), type="html",html.table.attributes="class='table table-striped table-hover table-bordered'", caption.placement="top", comment=FALSE, include.rownames=TRUE, include.colnames=TRUE)
    }

ProjectData_segment <- NEW_ProjectData
```
</div>
</div>
<br> <br>


```{r Fig2, echo=FALSE, comment=NA, results='asis', message=FALSE, echo=FALSE, fig.align='center', fig=TRUE}

if(ncol(NEW_ProjectData) >= 2) {
  df1  <- cbind(NEW_ProjectData[, 1], NEW_ProjectData[, 2])
  df1  <- as.data.frame(df1)
  sca1 <- gvisScatterChart(df1, options=list(legend="none",
                                             lineWidth=0, pointSize=8, hAxis.title="Derived Variable (Factor) 2",
                                             title="Data Visualization Using the top 2 Derived Attributes (Factors)", vAxis="{title:'Derived Variable (Factor) 1'}",
                                             hAxis="{title:'Derived Variable (Factor) 2'}", width=900, height=800))
  print(sca1,'chart')
  } else {
    df2  <- cbind(NEW_ProjectData[, 1], ProjectData[, 1])
    df2  <- as.data.frame(df2)
    sca2 <- gvisScatterChart(df2, options=list(legend="none",
                                               lineWidth=0, pointSize=12, hAxis.title="Derived Variable (Factor) 1",
                                               title="Only 1 Derived Variable: Using Initial Variable", vAxis="{title:'Derived Variable (Factor) 1'}",
                                               hAxis="{title:'Initial Variable (Factor) 1'}", width=900, height=800))
    print(sca2,'chart')
    }

```


Boats Segmentation: Cluster Analysis

========================================================


Business Decisions
---------------------------------------------------------

your text...



The Data
--------------------------------------------


Your text


```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
#ProjectData = data.matrix(ProjectData)
```

<br>

Here are the responses for the first `r min(max_data_report,nrow(ProjectData))` people:

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectData,2))

show_data = show_data[1:min(max_data_report,nrow(show_data)),]

row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>


Our Approach
---------------------------------------------------------------

#### Clustering and Segmentation in 9 steps

1. Confirm the data in metric 

2. Decide whether to scale or standardize the data

3. Decide which variables to use for clustering

4. Define similarity or dissimilarity measures between observations

5. Visualize Individual Attributes and  Pair-wise Distances between the Observations

6. Select the clustering method to use and decide how many clusters to have

7. Profile and interpret the clusters 

8. Assess the robustness of our clusters

Let's follow these steps.

#### Step 1. Confirm the data in metric 


In our case the data are metric, so we continue to the next step. Before doing so, we see the descriptive statistics of our data to get, as always, a better understanding of the data. 
Our data have the following descriptive statistics: 

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectData),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>



#### 2. Decide whether to scale or standardize the data


```{r results='asis'}
ProjectData_scaled <- apply(ProjectData,2, function(r) {if (sd(r)!=0) res=(r-mean(r))/sd(r) else res=0*r; res})

#ProjectData_scaled <- ProjectData

#print(ProjectData_scaled)

```

Notice now the summary statistics of the scaled dataset:

<br>

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}

show_data = data.frame(round(my_summary(ProjectData_scaled),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>

<br>
As expected all variables have mean 0 and standard deviation 1. 

While this is typically a necessary step, one has to always do it with care: some times you may want your analytics findings to be driven mainly by a few attributes that take large values; other times having attributes with different scales may imply something about those attributes. In many such cases one may choose to skip step 2 for some of the raw attributes.  

We decide to use the unscaled data for now. 


#### Step 3. Decide which variables to use for clustering

Line 204. Currently it used the factor scores. Need explanations.


#### Step 4. Define similarity or dissimilarity measures between observations

In our case we explore two distance metrics: the commonly used **Euclidean distance** as well as a simple one we define manually. 

The Euclidean distance between two observations (in our case, customers) is simply the square root of the average of the square difference between the attributes of the two observations (in our case, customers). For example, the distance of the first customer in our data from customers 2-5 (summarized above), using their responses to the 6 attitudinal questions is:

```{r include=FALSE, echo=FALSE, comment=NA, warning=FALSE, message=FALSE}
euclidean_pairwise <- as.matrix(dist(head(ProjectData_segment, 5), method="euclidean"))
euclidean_pairwise <- euclidean_pairwise*lower.tri(euclidean_pairwise) + euclidean_pairwise*diag(euclidean_pairwise) + 10e10*upper.tri(euclidean_pairwise)
euclidean_pairwise[euclidean_pairwise==10e10] <- NA
```

<div class="row">
<div class="col-md-4">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(euclidean_pairwise, caption="Pairwise Distances between the first 5 observations using The Euclidean Distance Metric", digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>

Notice for example that if we use, say, the Manhattan distance metric, these distances change as follows:

```{r include=FALSE, echo=FALSE, comment=NA, warning=FALSE, message=FALSE}
manhattan_pairwise <- as.matrix(dist(head(ProjectData_segment, 5), method="manhattan"))
manhattan_pairwise <- manhattan_pairwise*lower.tri(manhattan_pairwise) + manhattan_pairwise*diag(manhattan_pairwise) + 10e10*upper.tri(manhattan_pairwise)
manhattan_pairwise[manhattan_pairwise==10e10] <- NA
```

<div class="row">
<div class="col-md-4">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(manhattan_pairwise, caption="Pairwise Distances between the first 5 observations using The Manhattan Distance Metric", digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>

Let's now define our own distance metric, as an example. Let's say that the management team of the company believes that two customers are similar if they do not differ in their ratings of the attitudinal questions by more than 2 points. We can manually assign a distance of 1 for every question for which two customers gave an answer that differs by more than 2 points, and 0 otherwise. It is easy to write this distance function in R:

```{r ,results='asis'}
My_Distance_function<-function(x,y){sum(abs(x-y)>2)}

```

Here is how the pairwise distances between the respondents now look like.

```{r include=FALSE, echo=FALSE, comment=NA, warning=FALSE, fig.align='center', message=FALSE}
Manual_Pairwise=apply(head(ProjectData_segment,5),1,function(i) apply(head(ProjectData_segment,5),1,function(j) My_Distance_function(i,j) ))
Manual_Pairwise <- Manual_Pairwise * lower.tri(Manual_Pairwise) + Manual_Pairwise * diag(Manual_Pairwise) + 10e10*upper.tri(Manual_Pairwise)
Manual_Pairwise[Manual_Pairwise == 10e10] <- NA
```

<div class="row">
<div class="col-md-4">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
print(xtable(Manual_Pairwise, caption="Pairwise Distances between the first 5 observations using a simple manually defined Distance Metric", digits=1), type="html", html.table.attributes = "class='table table-striped table-hover table-bordered'", caption.placement="top", comment = FALSE, include.rownames = FALSE)
```
</div>
</div>

In general a lot of creative thinking and exploration should be spent in this step, and as always one may need to come back to this step even after finishing the complete segmentation process - multiple times. 

#### Step 5. Visualize Individual Attributes and  Pair-wise Distances between the Observations


For example, in our case we can see the histogram of, say, the first 4 variables:

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, echo=FALSE, error=FALSE, fig.align='center', results='asis'}

if (0){
  
  ProjectDataframe = data.frame(ProjectData_segment[,1:min(4,ncol(ProjectData_segment))])
  V1 <- ggplot() + geom_bar(aes(y = ..count.., x = V1),data=ProjectDataframe) + ylab(label = 'Frequency') + xlab(label = 'Histogram of Variable 1') + theme_grey()
  V2 <- ggplot() + geom_bar(aes (y = ..count.., x = V2),data=ProjectDataframe) + ylab(label = 'Frequency') + xlab(label = 'Histogram of Variable 2') + theme_grey()
  V3 <- ggplot() + geom_bar(aes (y = ..count.., x = V3),data=ProjectDataframe) + ylab(label = 'Frequency') + xlab(label = 'Histogram of Variable 3') + theme_grey()
  V4 <- ggplot() + geom_bar(aes (y = ..count.., x = V4),data=ProjectDataframe) + ylab(label = 'Frequency') + xlab(label = 'Histogram of Variable 4') + theme_grey()
  grid.arrange(V1, V2, V3, V4)
  
  }

```

or the histogram of all pairwise distances for the `r distance_used` distance:


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
Pairwise_Distances <- dist(ProjectData_segment, method = distance_used) 
hist(Pairwise_Distances, main = NULL, xlab="Histogram of all pairwise Distances between observtions", ylab="Frequency")
```

<blockquote> <p>
Visualization is very important for data analytics, as it can provide a first understanding of the data.
</p> </blockquote>

<br> 

#### Step 6. Select the clustering method to use and decide how many clusters to have

YOUR TEXT

In this note we will use two widely used methods: the **Kmeans Clustering Method**, and the **Hierarchical Clustering Method**. Like all clustering methods, these two also require that we have decided how to measure the distance/similarity between our observations.  Explaining how these methods work is beyond our scope. The only difference to highlight is that Kmeans requires the user to define how many segments to create, while Hierarchical Clustering does not. 

Let's fist use the **Hierarchial Clustering** method, as we do not know for now how many segments there are in our data. Hierarchical clustering is a  method that also helps us visualise how the data may be clustering together. It generates a plot called the **Dendrogram** which is often helpful for visualization - but should be used with care. For example, in this case the dendrogram, using the `r distance_used` distance metric from the earlier steps and the ``r hclust_method` hierarchical clustering option (see below as well as help(hclust) in R for more information), is as follows:

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
Hierarchical_Cluster_distances <- dist(ProjectData_segment, method=distance_used)
Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=hclust_method)
# Display dendogram
plot(Hierarchical_Cluster, main = NULL, sub=NULL, labels = 1:nrow(ProjectData_segment), xlab="Our Observations", cex.lab=1, cex.axis=1) 
# Draw dendogram with red borders around the 3 clusters
rect.hclust(Hierarchical_Cluster, k=numb_clusters_used, border="red") 
```

Note that we can draw as many clusters are we choose (e.g. in this case we chose `r numb_clusters_used` clusters) around the branches of the Dendrogram.


We can also plot the "distances" traveled before we need to merge any of the lower and smaller in size clusters into larger ones - the heights of the tree branches that link the clusters as we traverse the tree from its leaves to its root. If we have n observations, this plot has n-1 numbers. 


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, fig.align='center', results='asis'}
max <- nrow(ProjectData)
num <- max - 1
df1 <- cbind(as.data.frame(Hierarchical_Cluster$height[length(Hierarchical_Cluster$height):1]), c(1:num))
colnames(df1) <- c("distances","index")
Line <- gvisLineChart(as.data.frame(df1), xvar="index", yvar="distances", options=list(title='Distances plot', legend="right", width=900, height=600, hAxis="{title:'Number of Components', titleTextStyle:{color:'black'}}", vAxes="[{title:'Distances'}]", series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(Line,'chart')
```

For now let's consider the `r numb_clusters_used`-segments solution found by the Hierarchical Clustering method (using the `r distance_used` distance and the hclust option `r hclust_method`). We can also see the segment each observation (respondent in this case) belongs to for the first `r min(max_data_report,nrow(ProjectData))` people:


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
cluster_memberships_hclust <- as.vector(cutree(Hierarchical_Cluster, k=numb_clusters_used)) # cut tree into 3 clusters
cluster_ids_hclust=unique(cluster_memberships_hclust)

ProjectData_with_hclust_membership <- cbind(1:length(cluster_memberships_hclust),cluster_memberships_hclust)
colnames(ProjectData_with_hclust_membership)<-c("Observation Number","Cluster_Membership")
```


<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectData_with_hclust_membership,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>

**Using Kmean Clustering**

Here are the clusters our observations belong to when we select `r numb_clusters_used` clusters and the `r kmeans_method` kmeans method, for the first `r min(max_data_report,nrow(ProjectData))` people:


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
kmeans_clusters <- kmeans(ProjectData_segment,centers= numb_clusters_used, iter.max=1000, algorithm=kmeans_method)

ProjectData_with_kmeans_membership <- cbind(1:length(kmeans_clusters$cluster),kmeans_clusters$cluster)
colnames(ProjectData_with_kmeans_membership)<-c("Observation Number","Cluster_Membership")

## getting the cluster ids into Project Data

ProjectData <- cbind(ProjectData, kmeans_clusters$cluster)
colnames(ProjectData)[ncol(ProjectData)] <- "Cluster_Membership"
cluster_membership_test <- rep(0, nrow(ProjectData))
ProjectData <- cbind(ProjectData, cluster_membership_test)
colnames(ProjectData)[ncol(ProjectData)] <- "Cluster_test"

ProjectData[ProjectData[, 45] == Cluster_to_test , 46] <- 1

Probability_Threshold = Probability_Threshold/100 # make it between 0 and 1
dependent_variable = unique(sapply(dependent_variable,function(i) min(ncol(ProjectData), max(i,1))))
independent_variables = unique(sapply(independent_variables,function(i) min(ncol(ProjectData), max(i,1))))

if (length(unique(ProjectData[,dependent_variable])) !=2){
  cat("\n*****\n BE CAREFUL, THE DEPENDENT VARIABLE TAKES MORE THAN 2 VALUES...")
  cat("\nSplitting it around its median...\n*****\n ")
  new_dependent = ProjectData[,dependent_variable] >= median(ProjectData[,dependent_variable])
  ProjectData[,dependent_variable] <- 1*new_dependent
}

```

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectData_with_kmeans_membership,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>


#### Step 7. Profile and interpret the clusters 



YOUR TEXT...

In our case, assuming we decided we use the `r numb_clusters_used` segments found using Kmeans as outlined above (similar profiling can be done with the results of hierarchical clustering or other segmentation methods), we can see how the responses to our survey differ across segments. The average values of our data for the total population as well as within each customer segment are:


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
cluster_memberships_kmeans <- kmeans_clusters$cluster 
cluster_ids_kmeans <- unique(cluster_memberships_kmeans)
```

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
cluster_memberships <- cluster_memberships_hclust
cluster_ids <-  cluster_ids_hclust  
if (profile_with == "hclust"){
  cluster_memberships <- cluster_memberships_hclust
  cluster_ids <-  cluster_ids_hclust  
  }
if (profile_with == "kmeans"){
  cluster_memberships <- cluster_memberships_kmeans
  cluster_ids <-  cluster_ids_kmeans
  }

# SAVE THE DATA in the cluster file
NewData = matrix(cluster_memberships,ncol=1)
write.csv(NewData,file="cluster_file.csv")

population_average = matrix(apply(ProjectData_profile, 2, mean), ncol=1)
colnames(population_average) <- "Population"
Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_profile[(cluster_memberships==i), ], 2, mean))
if (ncol(ProjectData_profile) <2)
  Cluster_Profile_mean=t(Cluster_Profile_mean)
colnames(Cluster_Profile_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")
cluster.profile <- cbind (population_average,Cluster_Profile_mean)
```


<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
show_data = data.frame(round(cluster.profile,2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')

```
</div>
</div>

We can also "visualize" the segments using **snake plots** for each cluster. For example, we can plot the means of the profiling variables for each of our clusters to better visualize differences between segments. For better visualization we plot the standardized profiling variables.

```{r fig.width=6, fig.height=6, message=FALSE, echo=FALSE, fig.align='center', warning=FALSE, fig=TRUE}
ProjectData_scaled_profile = ProjectData_scaled[, profile_attributes_used,drop=F]

Cluster_Profile_standar_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_scaled_profile[(cluster_memberships==i), ,drop = F], 2, mean))
if (ncol(ProjectData_scaled_profile) < 2)
  Cluster_Profile_standar_mean = t(Cluster_Profile_standar_mean)
colnames(Cluster_Profile_standar_mean) <- paste("Segment", 1:length(cluster_ids), sep=" ")

custColors <- rainbow(12)[c(1, 2, 5, 9, 11, 12)]

plot(Cluster_Profile_standar_mean[, 1,drop=F], type="l", col=custColors[1], main="Snake plot for each cluster", ylab="mean of cluster", xlab="profiling variables (standardized)",ylim=c(min(Cluster_Profile_standar_mean),max(Cluster_Profile_standar_mean))) 
for(i in 2:ncol(Cluster_Profile_standar_mean))
  lines(Cluster_Profile_standar_mean[, i], col=custColors[i])
```

Can we see differences between the segments? Do the segments differ in terms of their average household income and in terms of how often they visit the mall? What else can we say about these segments?

YOUR TEXT...

We can also compare the averages of the profiling variables of each segment relative to the average of the variables across the whole population. This can also help us better understand whether  there are indeed clusters in our data (e.g. if all segments are much like the overall population, there may be no segments). For example, we can measure the ratios of the average for each cluster to the average of the population (e.g. avg(cluster)/avg(population)) and explore a matrix as the following one:

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
population_average_matrix <- population_average[,"Population",drop=F] %*% matrix(rep(1,ncol(Cluster_Profile_mean)),nrow=1)
cluster_profile_ratios <- (ifelse(population_average_matrix==0, 0,Cluster_Profile_mean/population_average_matrix))
colnames(cluster_profile_ratios) <- paste("Segment", 1:ncol(cluster_profile_ratios), sep=" ")
rownames(cluster_profile_ratios) <- colnames(ProjectData)[profile_attributes_used]

## printing the result in a clean-slate table
cat(renderHeatmapX(cluster_profile_ratios, border=1, center = 1, minvalue = heatmin))
```


YOUR TEXT: INTERPRETATION

#### Step 8. Assess the robustness of our clusters

We will skip this step for this exercise. However in practice this is a very important step. 


Intepretation, Decisions and Conclusion
----------------------------------------------------------------

YOUR TEXT HERE....




Boats Segmentation: Cluster Analysis

========================================================

**Team names**


Business Decisions
---------------------------------------------------------

your text...



The Data
--------------------------------------------


Your text...

With the aid of a market research firm, the boating company gathered various data about the boating market in the US through interviews with almost 3,000 boat owners and intenders. The data consisted, among others, of 29 attitudes towards boating, which respondents indicated on a 5-point scale. They are listed below. Other types of information had been collected, such as demographics as well as information about the boats, such as the length of the boat they owned, how they used their boats, and the price of the boats. 

After analyzing the survey data (using for example factor and cluster analysis), the company managers decided to only focus on a few purchase drivers which they thought were the most important ones. They decided to perform the classification and purchase drivers analysis using only the responses to the following questions:


Let's get the data and see it for a few customers. This is how the first `r min(max_data_report, nrow(ProjectData))` out of the total of `r nrow(ProjectData)` rows look:
<br>


<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectData,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>

Our Approach
---------------------------------------------------------------

There is *not a single best* process for classification. However, we have to start somewhere, so we will use the following process:

#### Classification in 6 steps


#### Step 1: Splitting the data into estimation and validation samples

You text...

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='hide'}

if (random_sampling){
  estimation_data_ids=sample.int(nrow(ProjectData),floor(estimation_data_percent*nrow(ProjectData)/100))
  non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
  validation_data_ids=non_estimation_data[sample.int(length(non_estimation_data), floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))]
  } else {
    estimation_data_ids=1:floor(estimation_data_percent*nrow(ProjectData)/100)
    non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
    validation_data_ids = (tail(estimation_data_ids,1)+1):(tail(estimation_data_ids,1) + floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))
    }

test_data_ids = setdiff(1:nrow(ProjectData), union(estimation_data_ids,validation_data_ids))

estimation_data=ProjectData[estimation_data_ids,]
validation_data=ProjectData[validation_data_ids,]
test_data=ProjectData[test_data_ids,]
```


In our case we use for example `r nrow(estimation_data)` observations in the estimation data, 
`r nrow(validation_data)` in the validation data, and `r nrow(test_data)` in the test data. 


#### Step 2: Setting up the dependent variable


Your text...

In our data the number of 0/1's in our estimation sample is as follows:
<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
class_percentages=matrix(c(sum(estimation_data[,dependent_variable]==1),sum(estimation_data[,dependent_variable]==0)), nrow=1); colnames(class_percentages)<-c("Class 1", "Class 0")
rownames(class_percentages)<-"# of Observations"
print(xtable(class_percentages ,caption="Number of Observations per class in the Estimation Sample", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```

</div>
</div>
while in the validation sample they are:

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
class_percentages=matrix(c(sum(validation_data[,dependent_variable]==1),sum(validation_data[,dependent_variable]==0)), nrow=1); colnames(class_percentages)<-c("Class 1", "Class 0")
rownames(class_percentages)<-"# of Observations"
print(xtable(class_percentages ,caption="Number of Observations per class in the Validation Sample", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)
```
</div>
</div>


#### Step 3: Make a preliminary assessment of the relative importance of the explanatory variables using visualization tools and simple descriptive statistics.


<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}  
show_data = data.frame(round(my_summary(estimation_data[estimation_data[,dependent_variable]==1,]),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}  
show_data = data.frame(round(my_summary(estimation_data[estimation_data[,dependent_variable]==0,]),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=400,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>


Your text

<center style="width=1048px;">
```{r echo=FALSE, message=FALSE, warning=FALSE,prompt=FALSE, results='asis',fig.height=10,fig.width=16}
par(mfrow=c(2,1))
DVvalues = unique(estimation_data[,dependent_variable])
x0 = estimation_data[which(estimation_data[,dependent_variable]==DVvalues[1]),independent_variables]
x1 = estimation_data[which(estimation_data[,dependent_variable]==DVvalues[2]),independent_variables]
colnames(x0) <- 1:ncol(x0)
colnames(x1) <- 1:ncol(x1)
boxplot(x0)
boxplot(x1)
```
</center>

Can you see which variables appear to be the most discrimatory ones?
Your text


#### Step 4: Estimate the classification model using the estimation data, and interpret the results.

Your text


```{r echo=FALSE, message=FALSE,warning=FALSE, prompt=FALSE, results='asis'}

# just name the variables numerically so that they look ok on the tree plots
independent_variables_nolabel = paste("IV", 1:length(independent_variables), sep="")

estimation_data_nolabel = cbind(estimation_data[,dependent_variable], estimation_data[,independent_variables])
colnames(estimation_data_nolabel)<- c(colnames(estimation_data)[dependent_variable],independent_variables_nolabel)

validation_data_nolabel = cbind(validation_data[,dependent_variable], validation_data[,independent_variables])
colnames(validation_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

test_data_nolabel = cbind(test_data[,dependent_variable], test_data[,independent_variables])
colnames(test_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

estimation_data_nolabel = data.frame(estimation_data_nolabel)
validation_data_nolabel = data.frame(validation_data_nolabel)
test_data_nolabel = data.frame(test_data_nolabel)

estimation_data = data.frame(estimation_data)
validation_data = data.frame(validation_data)
test_data = data.frame(test_data)

```


<center>
```{r echo=FALSE, message=FALSE, warning=FALSE,prompt=FALSE, results='asis'}
formula=paste(colnames(estimation_data)[dependent_variable],paste(Reduce(paste,sapply(head(independent_variables_nolabel,-1), function(i) paste(i,"+",sep=""))),tail(independent_variables_nolabel,1),sep=""),sep="~")
CART_tree<-rpart(formula, data= estimation_data_nolabel,method="class", control=CART_control)

fancyRpartPlot(CART_tree)
```
</center>


<center style="width=900px; height=800px;">
```{r echo=FALSE, message=FALSE,warning=FALSE, prompt=FALSE, results='asis',fig.width=14,fig.height=10}
CART_tree_large<-rpart(formula, data= estimation_data_nolabel,method="class", control=rpart.control(cp = 0.005))

fancyRpartPlot(CART_tree_large)
```
</center>

In our case, the probability our validation data belong to class 1 (e.g. the customer is likely to purchase a boat) for the first few validation data observations, using the first CART above, is:

```{r echo=FALSE, message=FALSE,warning=FALSE, results='asis'}
# Let's first calculate all probabilites for the estimation, validation, and test data
estimation_Probability_class1_tree<-predict(CART_tree, estimation_data_nolabel)[,2]
estimation_Probability_class1_tree_large<-predict(CART_tree_large, estimation_data_nolabel)[,2]

validation_Probability_class1_tree<-predict(CART_tree, validation_data_nolabel)[,2]
validation_Probability_class1_tree_large<-predict(CART_tree_large, validation_data_nolabel)[,2]

test_Probability_class1_tree<-predict(CART_tree, test_data_nolabel)[,2]
test_Probability_class1_tree_large<-predict(CART_tree_large, test_data_nolabel)[,2]


estimation_prediction_class_tree=1*as.vector(estimation_Probability_class1_tree > Probability_Threshold)
estimation_prediction_class_tree_large=1*as.vector(estimation_Probability_class1_tree_large > Probability_Threshold)

validation_prediction_class_tree=1*as.vector(validation_Probability_class1_tree > Probability_Threshold)
validation_prediction_class_tree_large=1*as.vector(validation_Probability_class1_tree_large > Probability_Threshold)

test_prediction_class_tree=1*as.vector(test_Probability_class1_tree > Probability_Threshold)
test_prediction_class_tree_large=1*as.vector(test_Probability_class1_tree_large > Probability_Threshold)

```
<br>
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}

Classification_Table=rbind(validation_data[,dependent_variable],validation_Probability_class1_tree)
rownames(Classification_Table)<-c("Actual Class","Probability of Class 1")
colnames(Classification_Table)<- paste("Obs", 1:ncol(Classification_Table), sep=" ")

Classification_Table_large=rbind(validation_data[,dependent_variable],validation_Probability_class1_tree)
rownames(Classification_Table_large)<-c("Actual Class","Probability of Class 1")
colnames(Classification_Table_large)<- paste("Obs", 1:ncol(Classification_Table_large), sep=" ")

show_data = data.frame(round(Classification_Table,2))
show_data = show_data[,1:min(max_data_report,ncol(show_data))]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Classification Table"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=140,allowHTML=TRUE,page='disable'))
print(m1,'chart')

```
<br>

In practice we need to select the **probability threshold** above which we consider an observation as "class 1": this is an important choice that we will discuss below. First we discuss another method widely used, namely logistic regression.

<br>
<br>

**Logistic Regression** is a method similar to linear regression except that the dependent variable can be discrete (e.g. 0 or 1). **Linear** logistic regression estimates the coefficients of a linear model using the selected independent variables while optimizing a classification criterion. For example, this is the logistic regression parameters for our data:


<br>
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}

formula_log=paste(colnames(estimation_data[,dependent_variable,drop=F]),paste(Reduce(paste,sapply(head(independent_variables,-1), function(i) paste(colnames(estimation_data)[i],"+",sep=""))),colnames(estimation_data)[tail(independent_variables,1)],sep=""),sep="~")

logreg_solution <- glm(formula_log, family=binomial(link="logit"),  data=estimation_data)

log_coefficients = round(summary(logreg_solution)$coefficients,1)
print(xtable(log_coefficients,caption="Logistic Regression: Estimated Coefficients" , digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)

```

In our case, the probability our validation data belong to class 1 (e.g. the customer is likely to purchase a boat) for the first few validation data observations, using the logistic regression above, is:

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# Let's get the probabilities for the 3 types of data again
estimation_Probability_class1_log<-predict(logreg_solution, type="response", newdata=estimation_data[,independent_variables])
validation_Probability_class1_log<-predict(logreg_solution, type="response", newdata=validation_data[,independent_variables])
test_Probability_class1_log<-predict(logreg_solution, type="response", newdata=test_data[,independent_variables])

estimation_prediction_class_log=1*as.vector(estimation_Probability_class1_log > Probability_Threshold)
validation_prediction_class_log=1*as.vector(validation_Probability_class1_log > Probability_Threshold)
test_prediction_class_log=1*as.vector(test_Probability_class1_log > Probability_Threshold)

```

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}

Classification_Table=rbind(validation_data[,dependent_variable],validation_Probability_class1_log)
rownames(Classification_Table)<-c("Actual Class","Probability of Class 1")
colnames(Classification_Table)<- paste("Obs", 1:ncol(Classification_Table), sep=" ")

show_data = data.frame(round(Classification_Table,2))
show_data = show_data[,1:min(max_data_report,ncol(show_data))]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Classification Table"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=140,allowHTML=TRUE,page='disable'))
print(m1,'chart')

```
<br>


Your text

From this table we can see the **key drivers** of the classification according to each of the methods we used here. 

<br>
<br>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
log_importance = tail(log_coefficients[,"z value", drop=F],-1) # remove the intercept
log_importance = log_importance/max(abs(log_importance))

tree_importance = CART_tree$variable.importance
tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree$variable.importance)))
tree_importance_final = rep(0,length(independent_variables))
tree_importance_final[tree_ordered_drivers] <- tree_importance
tree_importance_final <- tree_importance_final/max(abs(tree_importance_final))
tree_importance_final <- tree_importance_final*sign(log_importance)

large_tree_importance = CART_tree_large$variable.importance
large_tree_ordered_drivers = as.numeric(gsub("\\IV"," ",names(CART_tree_large$variable.importance)))
large_tree_importance_final = rep(0,length(independent_variables))
large_tree_importance_final[large_tree_ordered_drivers] <- large_tree_importance
large_tree_importance_final <- large_tree_importance_final/max(abs(large_tree_importance_final))
large_tree_importance_final <- large_tree_importance_final*sign(log_importance)

Importance_table <- cbind(tree_importance_final,large_tree_importance_final, log_importance)
colnames(Importance_table) <- c("CART 1", "CART 2", "Logistic Regr.")
rownames(Importance_table) <- rownames(log_importance)
## printing the result in a clean-slate table
cat(renderHeatmapX(Importance_table, border=1, center = 0, minvalue = 0))
```

<br>
<br>

#### Step 5: Assess the accuracy of classification in the first validation sample

Your text

### 1.  **Hit ratio** 

These are as follows for the probability threshold  `r Probability_Threshold*100`% for the validation data:

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE,warning=FALSE,comment=NA,prompt=FALSE, results='asis'}
validation_actual=validation_data[,dependent_variable]
validation_predictions = rbind(validation_prediction_class_tree,
                               validation_prediction_class_tree_large,
                               validation_prediction_class_log)
validation_hit_rates = rbind(
  100*sum(validation_prediction_class_tree==validation_actual)/length(validation_actual), 
  100*sum(validation_prediction_class_tree_large==validation_actual)/length(validation_actual), 
  100*sum(validation_prediction_class_log==validation_actual)/length(validation_actual)
  )
colnames(validation_hit_rates) <- "Hit Ratio"
rownames(validation_hit_rates) <- c("First CART", "Second CART", "Logistic Regression")

print(xtable(validation_hit_rates ,caption="Validation Data Hit Ratios for different classifiers tested", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)

```
</div>
</div>


while for the estimation data the hit rates are:
<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
estimation_actual=estimation_data[,dependent_variable]
estimation_predictions = rbind(estimation_prediction_class_tree,
                               estimation_prediction_class_tree_large,
                               estimation_prediction_class_log)
estimation_hit_rates = rbind(
  100*sum(estimation_prediction_class_tree==estimation_actual)/length(estimation_actual), 
  100*sum(estimation_prediction_class_tree_large==estimation_actual)/length(estimation_actual), 
  100*sum(estimation_prediction_class_log==estimation_actual)/length(estimation_actual)
  )
colnames(estimation_hit_rates) <- "Hit Ratio"
rownames(estimation_hit_rates) <- c("First CART", "Second CART", "Logistic Regression")

print(xtable(estimation_hit_rates ,caption="Estimation Data Hit Ratios for different classifiers tested", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)

```
</div>
</div>

**Why are the performances on the estimation and validation data different? How different can they possibly be? What does this diffference depend on?** Is the Validation Data Hit Rate satisfactory? Which classifier should we use? What should be the benchmark against which to compare the hit rate? 

YOUR TEXT/INTERPRETATION

### 2. **Confusion matrix**

YOUR TEXT/INTERPRETATION

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
validation_prediction_best = validation_predictions[which.max(validation_hit_rates),]
conf_matrix = matrix(rep(0,4),ncol=2)
conf_matrix[1,1]<- 100*sum(validation_prediction_best*validation_data[,dependent_variable])/sum(validation_data[,dependent_variable])
conf_matrix[1,2]<- 100*sum((!validation_prediction_best)*validation_data[,dependent_variable])/sum(validation_data[,dependent_variable])
conf_matrix[2,1]<- 100*sum((!validation_prediction_best)*(!validation_data[,dependent_variable]))/sum((!validation_data[,dependent_variable]))
conf_matrix[2,2]<- 100*sum((validation_prediction_best)*(!validation_data[,dependent_variable]))/sum((!validation_data[,dependent_variable]))
conf_matrix = round(conf_matrix,2)

colnames(conf_matrix) <- c("Predicted 1", "Predicted 0")
rownames(conf_matrix) <- c("Actual 1", "Actual 0")

print(xtable(conf_matrix ,caption="Confusion Matrix for Validation data", digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE)
```

</div>
</div>
<br>

Note that the percentages add up to 100% for each row: can you see why? Moreover, a "good" confusion matrix should have large diagonal values and small off-diagonal oens: you see why?

YOUR TEXT/INTERPRETATION